
```{r}
anchor_items <- read_rds(here("verbalanalysis/03-anchor-items.rds"))
R.utils::sourceDirectory(here("R"))

emlg <-
    coef_1f(anchor_items$intuitive_mod[[1]])$items %>%
    select(item, ref_easy = a_ref_easy, foc_easy = b_foc_easy) %>%
    mutate(difference_in_easy = ref_easy - foc_easy) %>%
    left_join(item_map %>% select(item, full))

gini_results <- emlg_to_gini_l1(emlg, lo = -2, hi = 2)

min_bc <- read_rds(here("verbalanalysis/04-min-bc.rds"))
```

### Maximizing the Gini index (MAXGI)

@strobl2018anchor suggest using the Gini index [@gini1912variabilita] to select the anchor point. In general, the Gini index "takes high values if, for example, a small minority of persons has a lot of wealth while the vast majority has very little" [@strobl2018anchor, p. 7]. The Gini index is frequently used to measure the inequality of wealth distribution in a country; for example, South Africa typically has the highest Gini index of all measured countries, meaning that wealth is distributed unevenly [@chitiga2015income].

Here, DIF effectively takes the role of wealth; @strobl2018anchor select $\mu^{\star\text{foc}}$ by maximizing the Gini index of $\tilde{\mathbf{d}}$ (thus our abbreviation MAXGI). As a result, models where fewer items are responsible for the majority of deviations from the mean group difference in performance are preferred. Effectively, MAXGI selects the anchor point so that the majority of items having little to no DIF and a small subset of items thus have larger amounts of DIF. Denoting a function that calculates the Gini index from a vector of non-negative elements as  $G(\mathbf{x})$, MAXGI sets
\begin{equation}
\mu^{\star\text{foc}} = \argmax_{\mu^\text{foc}} G(|\mu^\text{foc} + \tilde{\mathbf{d}}|)
\end{equation}
where $\mu^\text{foc} \in (-\infty, \infty)$ and $\mu^\text{foc}$ is added to each element of $\tilde{\mathbf{d}}$. For the verbal aggression data, Figure \ref{fig:thegini} shows the gini index at a variety of possible $\mu^\text{foc}$ values. The result of MAXGI is $\mu^{\star\text{foc}} = `r round(gini_results$gini_anchor_points, 3)`$.^[There is also a local maximum at $\mu^{\text{male}} \approx -0.75$, which corresponds to the items that females give affirmative responses to. This illustrates that—as @strobl2018anchor point out—the search path is informative.]  

```{r thegini, fig.height = 2.5, fig.cap = 'The search path of MAXGI the verbal aggression data. $\\mu^\\text{female}$ is fixed to 0 and the goal is to identify the value of $\\mu^\\text{male}$ that maximizes the gini coefficient. Maximizing the gini coefficient leads to a small minority of items containing DIF.'}

gini_results$grid %>%
    ggplot(aes(x = -anchor_point, y = gini)) +
    geom_point() +
    labs(
        x = latex2exp::TeX("$\\mu^{male}$"),
        y = latex2exp::TeX("$G(|\\mu^{male} + \\tilde{\\mathbf{d}}|)$")
    ) +
    geom_point(
        data = gini_results$grid %>% filter(gini == max(gini)),
        color = "blue",
        size = 3
    ) +
    geom_vline(
        xintercept = -pull(gini_results$grid %>% filter(gini == max(gini)), anchor_point),
        linetype = "dashed"
    ) +
    scale_x_continuous(breaks = -8:8 / 4)
```

### Minimizing Between Curves (MINBC)

We propose an alternative to MAXGI that selects the anchor point by minimizing the total amount of DIF on the test. Accordingly, we name this method "minimizing between curves" (MINBC). MINBC is inspired by Raju's area method [@raju1988area], which measures the amount of DIF by calculating the area between the item characteristic curves, the function that maps the student's ability to their probability of correct response, of the two groups:
\begin{equation}
\text{Area Between Curves} = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, b_j^{\text{foc}})| d\theta
\end{equation}
Raju's area method has been cited as one of the most commonly used IRT-based DIF detection methods [@magis2011generalized]. However, Raju's area method is not an AGI method because the item characteristic curves first need to be linked. An additional weakness is that the area is typically unweighted, so all values of $\theta$ within a range—commonly from -6 to 6—matter equally, despite some being much more realistic than others. 
MINBC adapts Raju's area method into an AGI method by providing an algorithm for selecting the anchor point. MINBC chooses the anchor point that minimizes the total weighted area between each group's item characteristic curves. It is implemented as follows. We denote $m(\mu^\text{foc})$ to be a function that takes $\mu^\text{foc}$ as input and estimates $\hat b_j^\text{foc}$ by fitting a unidimensional Rasch model. The amount of DIF on each item is calculated as
\begin{equation}
\text{DIF}_j(\mu^\text{foc}) = \int |\text{Pr}(y_j = 1| \theta, b_j^{\text{ref}}) - \text{Pr}(y_j = 1| \theta, m_j(\mu^\text{foc}))| g(\theta)d\theta
\end{equation}
where $g(\theta)$ is a weighting function such that $\int g(\theta)d\theta = 1$. $\text{Total DIF}(\mu^\text{foc})$ is a function where the input is $\mu^\text{foc}$ and the output is the total amount of DIF on the test:
\begin{equation}
\text{Total DIF}(\mu^\text{foc}) = \sum_{j} \text{DIF}_j(\mu^\text{foc}).
\end{equation}
MINBC sets the anchor point to be the value of $\mu^{\star\text{foc}}$ that minimizes the total amount of DIF on the test:
\begin{equation}
\mu^{\star\text{foc}} = \argmin_{\mu^\text{foc}} \text{Total DIF}(\mu^\text{foc}).
\end{equation}
MINBC is inspired in part by @chalmers2016might, who use the difference between test characteristic curves weighted by $g(\theta)$ as a measure of differential test functioning (DTF). The selection of $g(\theta)$ results in the relative weighting of $\theta$ values. Chalmers, Counsell, and Flora do not discuss how to choose $g(\theta)$ and in their empirical examples use $g(\theta)$ uniform for $-6 \le \theta \le 6$, which may be suboptimal in some cases. It might seem intuitive to choose $g(\theta) \sim N(0, 1)$ because $\mu^\text{female} = 0$, but this choice doesn't take into account the ability distribution of the focal group. For example, if $\mu^\text{foc} = 3$, we would want to prioritize higher $\theta$ values. Based on this logic, we set $g(\theta)$ to be the weighted average of the reference and focal group ability probability density functions:
\begin{equation}
g(\theta) = p^\text{ref} \cdot N(\mu^{\text{ref}}, \sigma^{\text{ref}^2}) + (1 - p^\text{ref}) \cdot N(\mu^{\text{foc}}, \sigma^{\text{foc}^2}).
\end{equation}

To understand MINBC, imagine a scenario in which the data-generating process is $\mu^{\text{foc}} = \mu^{\text{ref}}$ and $d_j = 0 \forall j$, so that the groups have equal ability and there is no DIF. The Fundamental DIF Identification Problem is that there are an infinite number of models with the same likelihood from which to choose. For example, we could correctly assume that the focal group has the same ability as the reference group and fix $\mu^{\star\text{foc}} = 0$. The model would then estimate $\hat d_j \approx 0 \forall j$, and we would correctly conclude the groups have the same ability and there is no DIF. Alternatively, we could assume that the focal group has $\mu^{\star\text{foc}} = 3$. The model would then compensate by finding $\hat d_j \approx -3 \forall j$, and we would incorrectly conclude that the focal group is high ability, but every item contains DIF against them. Both of these models have the same likelihood, so how should one choose which model to believe? MINBC chooses the model with the least amount of total DIF, as measured by the total weighted area between the item characteristic curves. As a result, the likelihood tie is broken by preferring to explain differences in performance across groups by ability differences (as opposed to DIF). Figure \ref{fig:minbc} shows Total DIF at a variety of possible values for $\mu^\text{male}$. The result is $\hat \mu^\text{male} \approx 0.34$.

```{r minbc, fig.height = 2.5, fig.cap = 'The search path of MINBC for the verbal aggression data. $\\mu^\\text{female}$ is fixed to 0 and the goal is to identify the value of $\\mu^\\text{male}$ that minimizes the total area between item characteristic curves. As a result, the total amount of DIF on the test is minimized and as much performance difference as possible is explained by group ability differences.'}
min_bc$grid %>%
    ggplot(aes(x = anchor_point, y = total_between_curves)) +
    geom_path(size = 2) +
    labs(
        x = latex2exp::TeX("$\\mu^{male}$"),
        y = latex2exp::TeX("$Total \\; DIF(\\mu^{male})$")
    ) +
    geom_vline(
        xintercept = min_bc$between_curves_anchor_points,
        linetype = "dashed"
    ) +
    geom_point(
        data = tibble(anchor_point = min_bc$between_curves_anchor_points, total_between_curves = 1.96),
        color = "blue",
        size = 3
    ) +
    scale_x_continuous(breaks = -8:8 / 4)
```
