In this section, we summarize existing AGI methods, propose extensions, and demonstrate their use on the verbal aggression data. So as to offer a coherent framework, we sometimes edit names of existing methods.^[We recognize that others have done the same [e.g., @kopf2015framework], and that we risk contributing to a proliferation of names.] Recall that the challenge presented by the Fundamental DIF Identification Problem for any AGI method is to disentangle estimation of $\hat\mu^\text{foc} - \mu^\text{ref} = \hat\mu^\text{foc} - 0 = \hat\mu^\text{foc}$, the difference in mean abilities, from the estimation of $\hat d_j = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$, the difference in item easiness. There are two classes of AGI methods. The most common, identification of anchor items, selects a group of items that are assumed to be DIF-free. These anchor items identify the model, thereby allowing for the estimation of $\hat\mu^\text{foc}$; the remaining items can be tested for DIF using an LRT. The other class involves identification of an anchor point (i.e., directly setting $\mu^{\text{foc}}$ to some value) [@strobl2018anchor]. This anchor point identifies the model, and all items can be tested for DIF using an LRT. We begin by describing anchor item methods.

## Anchor Item Methods

```{r child = "paper_3a_anchor_items.Rmd"}
```

## Anchor Point Methods

The previously discussed AGI methods identify the model by selecting a set of anchor items that are assumed to be DIF-free. We now discuss the second type of AGI methods: The use of an anchor point [@strobl2018anchor]. An anchor point directly sets the difference in mean abilities to some value, $\mu^{\star\text{foc}}$. Anchor point methods have the advantage of not requiring the assumption that any particular item is DIF-free, and, therefore, allowing all items to be tested for DIF. We describe two methods for selecting the anchor point.

```{r child = "paper_3b_anchor_points.Rmd"}
```

## Comparing Methods 

We described a variety of AGI methods. Anchor item methods—AOAA, AOAA-AS, and AOAA-OAT—address the Fundamental DIF Identification Problem by assuming a set of items do not contain DIF. Anchor point methods—MAXGI and MINBC—implement a rule to directly find group latent trait means. 

Each of these methods implemented an (arguably) reasonable algorithm to identify the final item response model. These final item response models each contain an estimate of the male-female group difference in verbal aggression, which are shown as vertical lines in Figure \ref{fig:summary}. Figure \ref{fig:summary} reveals that the conclusions arrived at vary significantly by DIF detection method. If the analyst were to use AOAA-OAT, they would find that males and females have nearly the same verbal aggression on average. On the other hand, if they were to use MINBC, they would find that males have significantly greater verbal aggression than females. A natural question to ask is, which AIG method is right? Instead of one method being best—which DIF detection method research has searched long and hard for—the variety of results is a consequence of the Fundamental DIF Identification Problem being fraught for the verbal aggression data as shown by the geometry of the GLIMMER in Figure \ref{fig:emmilg}. This is evidence that none of the AGI methods should be fully trusted.

```{r summary, fig.height = 4, fig.cap = 'Results by DIF detection method. The male group mean verbal aggression, $\\hat \\mu^\\text{male}$, according to each method is graphed as a vertical line, which are superimposed over the performance difference, $\\tilde {d_j}$, for each item. The scale is set by fixing $\\hat \\mu^\\text{female} = 0$.'}

anchor_items <- read_rds(here("verbalanalysis/03-anchor-items.rds"))
item_map <- read_rds(here("verbalanalysis/02-itemmap.rds"))

lines2 <- read_rds(here("verbalanalysis/07-lines2.rds"))
    
anchor_items$intuitive_mod[[1]] %>%
    coef_1f() %>%
    .$items %>%
    mutate(diff =  b_foc_easy - a_ref_easy) %>%
    left_join(read_rds(here("verbalanalysis/02labels.rds")) %>% rename(item = item_num)) %>%
    ggplot(aes(x = diff, y = full)) +
    geom_point() +
    geom_vline(
        data = lines2,
        aes(xintercept = Mean, linetype = Method, color = Method),
        size = 0.7
    ) +
    theme_minimal(base_size = 10) +
    theme(legend.position="bottom") +
    labs(
        x = latex2exp::TeX("$\\tilde{d_j} = \\tilde{b_j}^{male} - \\tilde{b_j}^{female}$"),
        y = "",
        color = "",
        linetype = ""
    ) +
    scale_x_continuous(breaks = -8:8/4)
```
